---
title: "Esp_football database analysis"
author: "Hajdú Márk"
date: '2022 03 21 '
output:
  pdf_document: default
  word_document: default
---

## Setting up my environment
Setting up R environment by loading the `tidyverse` and `readxl` packages and setting the working directory

```{r setup, message=FALSE, warning=FALSE}

library( "readxl" )
library("tidyverse")
setwd("C:/Users/Márk/Documents/okonometria2")
```

## Data importing
Importing the excel file which contains the database

```{r data import, echo=TRUE}
file<-"c:/Users/Márk/Documents/okonometriajo/ESPfootball.xlsx"
Esp_football<- read_excel( "ESPfootball.xlsx", sheet= 1)
```

## Data exploration
Previewing the data

```{r preview, echo=TRUE}
str(Esp_football)

```
## Data cleaning
### Changing data types
The Position and the Price are in a bad data type, I change them in order to do the analysis
```{r changing data type}
Esp_football$Position<-as.factor(Esp_football$Position)
Esp_football$PricemillionEuro<-as.numeric(Esp_football$PricemillionEuro)

```

### Basic statistics about the data
```{r summary}
summary(Esp_football)
```

### Filtering out the wrong data
There are rows that contain unreal, very big vaules, these are probably wrong data. Now I will filter them out


```{r filtering}
Esp_football_filtered<- Esp_football %>% filter(PricemillionEuro<101 &
                          Esp_football$Passes<100)
```

### Basic statistics about the filtered database

```{r summary 2}
summary(Esp_football_filtered)
```

Now there are not any very big outliers, I can start the analysis

## Linear regression
I will make a linear regression, the result variable will be the PriceMillionEuro
```{r}
model_1 <- lm(PricemillionEuro ~ Position + Age + YellowCards + RedCards + Goalsscored + 
                Foulscommitted + Shots + Shotsontarget + Assists + Passes, data = Esp_football_filtered)
summary(model_1)
```

There are a lot of insignificant variables, that means these variables are not in correlation with the result variable.
I will apply the backward elimination and I will only show the final model

## Final model
```{r}
model_8 <- lm(PricemillionEuro ~ Position + Age
              + Passes, data = Esp_football_filtered)
summary(model_8)
```

All variables are significant, except the midfielder dummy. I leave it in the model, because I do not want to change the reference category. The model has a low R2, it needs a lot of improvement and new variables.

## Data visualization

### The relationship between the age and the market value

```{r}

library(ggplot2)
ggplot(data = Esp_football_filtered, mapping = aes(x = Age, y = PricemillionEuro)) +
  geom_point(color="blue")+
  labs(
    title = "The relationship between the age and the price",
    subtitle = "There is a parabola relationship ",
    x = "Age",
    y = "Players' market value (in Million EUR)")
```

It would be proper to include the square of the age variable also and it would model that parabola line

### The relationship between the assists and the market value

```{r}
ggplot(data = Esp_football_filtered, mapping = aes(x = Assists, y = PricemillionEuro,color=Position)) + facet_wrap(~Position)+
  geom_point()+
  labs(
    title = "The relationship between the assists and the price",
    subtitle = "Grouped by position",
    x = "Assists(per match)",
    y = "Players' market value (in Million EUR)")
```

There is a surprising negative correlation betewwen the assist and the market value among the midfielders. There is a positive correlation among the forwards. Maybe that is the reason why the assists variable is not significant.



### The relationship between the goals and the market value

```{r}

ggplot(data = Esp_football_filtered, mapping = aes(x = Goalsscored, y = PricemillionEuro,color=Position)) + facet_wrap(~Position)+
  geom_point()+
  labs(
    title = "The relationship between the goals and the price",
    subtitle = "Grouped by position",
    x = "Goals(per match)",
    y = "Players' market value (in Million EUR)")
```

The goals variable is definitely a significant variable among the forwards. It would be worth to group the players by position and do separate linear regressions.

### Histogram 
```{r}
hist(Esp_football$RedCards,
     
     main="Distribution of players according to the red cards",
     xlab="Red cards(per match)",
     ylab = "Number of players",
     col = "lightblue")
```

The most players never get a red card. So every player will be an outlier if they get a red card.

## Key takeaways

* This final model is a good base model for further analysis
* According to the final model, the position,age and passes are the most important variables if we analyze the players' marekt value
* It would be worth to analyze the players separately, grouped by position
* The square of the age and other new variables might increase the explanatory power


